### 2020-12-24 paper/sex_selection_spacing-ver5.tex
Yet another abstract

\noindent India has experienced an increasingly male-biased sex ratio at birth following
the introduction of prenatal sex determination techniques. 
I examine three questions.
First, how did birth spacing change with this and other secular developments? 
Second, did changes in birth spacing make us overestimate the decline in cohort fertility? 
Finally, did the advent of sex-selective abortions change the relationship between birth 
spacing and infant mortality? 
Well-educated Hindu women with no sons had the most substantial lengthening of birth 
intervals and the most male-biased sex ratios, likely from sex-selective abortions. 
Consequently, some women with no sons now have longer birth intervals than those with sons, 
reversing India's traditional spacing pattern.
Less-educated women continued short spacing when they have girls, with only limited
evidence of male-biased sex ratios. 
Because of the overall lengthening in spacing, period fertility rates substantially 
overestimated how fast cohort fertility fell. 
Furthermore, predicted cohort fertility remains above replacement for all but the 
best-educated urban women. 
If lengthening of birth intervals occurs from repeated abortions, the associated
short pregnancy spacing may counteract any positive effects of longer birth spacing. 
There is no evidence of this effect on infant mortality. 
Finally, judging from sex ratios, sex-selective abortion use is not declining.

>>>>

% \noindent India has experienced an increasingly male-biased sex ratio at birth following
% the introduction of prenatal sex determination techniques. 
% I examine three questions.
% First, how did birth spacing change with this and other secular developments? 
% Second, did changes in birth spacing make us overestimate the decline in cohort fertility? 
% Finally, did the advent of sex-selective abortions change the relationship between birth 
% spacing and infant mortality? 
% Well-educated Hindu women with no sons had the most substantial lengthening of birth 
% intervals and the most male-biased sex ratios, likely from sex-selective abortions. 
% Consequently, some women with no sons now have longer birth intervals than those with sons, 
% reversing India's traditional spacing pattern.
% Less-educated women continued short spacing when they have girls, with only limited
% evidence of male-biased sex ratios. 
% Because of the overall lengthening in spacing, period fertility rates substantially 
% overestimated how fast cohort fertility fell. 
% Furthermore, predicted cohort fertility remains above replacement for all but the 
% best-educated urban women. 
% If lengthening of birth intervals occurs from repeated abortions, the associated
% short pregnancy spacing may counteract any positive effects of longer birth spacing. 
% There is no evidence of this effect on infant mortality. 
% Finally, judging from sex ratios, sex-selective abortion use is not declining.

\noindent Following the arrival of prenatal sex-determination technologies in the mid-1980s, 
India has experienced an increasingly male-biased sex ratio at birth, presumably from sex-selective 
abortions. 
Abortions lengthen birth intervals, but we know little about how birth spacing has changed or the 
effects of these changes. 
This paper shows that the length of birth intervals has increased overall, but well-educated women 
with no sons had the most substantial lengthening---and the most male-biased sex ratios. 
Furthermore, most of the changes took place immediately after the introduction of sex-selective abortions. 
Consequently, some women without sons now have longer birth intervals than those with sons, reversing 
India's traditional spacing pattern. 
Less-educated women continue short birth spacing when they have no sons, with only limited evidence 
of male-biased sex ratios. 
Because of the rapid lengthening of birth intervals, period fertility rates substantially overestimated 
how fast cohort fertility fell. 
Moreover, predicted cohort fertility is still 10\%--20\% above the period fertility rate. 
If lengthening of birth intervals occurs from repeated abortions, the associated short pregnancy spacing 
may counteract any positive effects of longer birth spacing. 
There is, however, no evidence of this effect on infant mortality. 
Finally, judging from sex ratios, sex-selective abortion use is not declining.
